---
title: "Statistical Inference"
author: "Kaveri Chatra"
output:
  html_document: default
  pdf_document: default
--- 

## Part 1: Investigating exponetial distribution and comparing with Central Limit Theorem

### Loading Libraries

```{r}
library("data.table")
library("ggplot2")
library("datasets")
library("rcompanion")
```

```{r, echo=TRUE}

set.seed(1)
lambda <- 0.2
sim<- 1000
n <- 40


exp_sim <- replicate(sim, rexp(n, lambda))
means_exp <- apply(exp_sim, 2, mean)
```

### 1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.

```{r, echo=TRUE}
sample_mean <- mean(exp_sim)
theoretical_mean <-1/lambda
hist(means_exp, xlab = "sample means", main = "Exponential Function")
abline(v = theoretical_mean, col = "orange", lwd = 4)
abline(v = sample_mean, col = "black")
```
From the graph it can be noticed that sample mean is very close to theoretical mean

### 2. Show how variable it is and compare it to the theoretical variance of the distribution.. 
Standard deviation
```{r, echo=TRUE}

sample_std <- sd(means_exp)
#For exp. distribution mean = std
theoretical_std <- 1/lambda/sqrt(n)
theoretical_std
sample_std
```
Variance
```{r, echo=TRUE}
sample_var <- var(means_exp)
theoretical_var <- 1/lambda^2/n
theoretical_var 
sample_var

```

### 3. Show that the distribution is approximately normal.
From the Central Limit Theorem, the distribution of averages is often normal, even if the distribution that the data is being sampled from is non-normal.

```{r, echo=TRUE}
par(mfrow = c(1,1))
hist(means_exp,breaks=n,prob=T,xlab = "means",ylab="density")
#curve(dnorm(x, 0, 1), -3, 3, col = 'blue',add=T) 
#lines(density(scale(means_exp)), col = 'red')
```

## Part 2:  Basic Inferential Data Analysis

### Loading and performing some exploratory data analysis
```{r, echo=TRUE}

head(ToothGrowth)
dim(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)
unique(ToothGrowth$dose)
```
#### Plots
```{r, echo = TRUE}
g <- ggplot(ToothGrowth, aes(len, supp, color = dose)) + geom_point()
print(g)
```

```{r, echo = TRUE}
g <- ggplot(ToothGrowth, aes(factor(dose), len, color = supp)) + geom_boxplot() + facet_grid(.~dose)
print(g)
```


### Using confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose

#### Confidence interval
``` {r, echo = TRUE}
groupwiseMean(len ~ supp, data = ToothGrowth, conf = 0.95, digits = 3)
groupwiseMean(len ~ factor(dose), data = ToothGrowth, conf = 0.95, digits = 3)
groupwiseMean(len ~ factor(dose) + factor(supp), data = ToothGrowth, conf = 0.95, digits = 3)
```
#### T test 
Tooth length corresponding to  OJ vs VC supplement. Null hypothesis being means of both are same and Alternate hypothesis being mean of  tooth growth corresponding to OJ is greater than that of VC supplements.
```{r, echo = TRUE}
OJ <- ToothGrowth$len[ToothGrowth$supp == 'OJ']
VC = ToothGrowth$len[ToothGrowth$supp == 'VC']
t.test(OJ, VC, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = 0.95)
```
Tooth Growth by dose. Dose is divided into 3 groups, group1 = 0.5, group2 = 1.0, group3 = 2.0
```{r, echo = TRUE}
group1 <- ToothGrowth$len[ToothGrowth$dose == 0.5]
group2 <- ToothGrowth$len[ToothGrowth$dose == 1.0]
group3 <- ToothGrowth$len[ToothGrowth$dose == 2.0]

t.test(group1, group2, alternative = "less", paired = FALSE, var.equal = FALSE, conf.level = 0.95)
t.test(group2, group3, alternative = "less", paired = FALSE, var.equal = FALSE, conf.level = 0.95)

```
```{r, echo=TRUE}
OJtwomg <- ToothGrowth$len[ToothGrowth$supp == 'OJ' & ToothGrowth$dose == 2.0]
VCtwomg <- ToothGrowth$len[ToothGrowth$supp == 'VC' & ToothGrowth$dose == 2.0]

t.test(OJtwomg, VCtwomg, alternative = "two.sided", paired = FALSE, var.equal = FALSE, conf.level = 0.95)
```
### Conclusions
From the result of the T tests and confidence intervals, it can be concluded that:
   
    1.  The supplement OJ had more impact than supplement VC on tooth growth.
    2.  Higher the dose more is the growth.
    3.  The impact on tooth growth 2.0 mg dose with VC supplement vs 2.0 mg dose with OJ supplement can't be determined.